An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-23 Revision/response.tex

\textbf{Regarding quantization time in index build:}

Of all the evaluated algorithms, only Faiss supports quantization. However, in our experiments, we used the IVF index structure of Faiss without employing quantization. Therefore, no quantization time was included in the reported index build times.

\textbf{Regarding Figure 7 and the exclusion of Vamana:}

We appreciate the observation of reviewer. Figure 7 presents experiments on multi-attribute construction and search, using Boolean AND logic. The Vamana-based methods (FilteredVamana and StitchedVamana) only support Boolean OR logic for multi-attribute search, which is why they were not included. We will clarify this point in the revised version of the paper.
Please see the section 4.2.2 on Multi-Attribute Construction and Multi-Attribute Query.

\textcolor{blue}{"Note that FilterDiskANN only supports Boolean OR logic and does not support Boolean AND. Therefore, its two Vamana-based methods are excluded from the multi-attribute construction and search experiments."}

%Thanks to the reviewer for suggesting that there is no explanation in Figure 7 why Vamana is not included. Figure 7 is an experiment on multi-label construction and search, and the Boolean logic of multi-label search is AND. The two algorithms using Vamana, FilteredVamana and StitchedVamana, only support Boolean logic OR for multi-label search, so we did not include them. We will add an explanation of this in the paper.

### 2025-07-23 VLDB2026/main.tex
Update main.tex
1

	%Similar to the evaluation of \textit{Attribute Filtering} algorithms, 
	Similarly, we also evaluate \textcolor{orange}{4} key metrics mentioned above in \textit{Range Filtering} algorithms. \textcolor{orange}{For the YT-Audio dataset with synthetic attributes and the YT-Audio dataset with real attributes, we can found that the data for index construction is basically the same, indicating that the authenticity of the attribute has almost no effect on index construction.
	However, the WIT dataset with synthetic labels is 2048-dimensional, while the WIT-Real dataset with real labels is 512-dimensional, so the dimension will cause a big difference.}

@@

	\textit{\textbf{Index size.}} 
	As shown in Figure~\ref{fig:rangeFilter_index_size_mb}, Among all algorithms, Faiss has the smallest index size. Because IVF only adds centroid data and partition information to the vector dataset, the IVF index is only slightly larger than the vector dataset.
	On low-dimensional datasets (Deep, YT-Audio), PASE requires the most storage. Its implementation likely wastes space due to alignment padding and inefficient encoding. %(e.g., Base64).

	SeRF maintains a compact index for low-dimensional data. However, for high-dimensional dataset (WIT), SeRF stores more edges and dynamic range data, significantly increasing its size.

	\textit{\textbf{Build peak memory.}}
	As Figure~\ref{fig:rangeFilter_build_memory_mb} shows, Faiss uses the least memory on low-dimensional datasets. It stores minimal temporary data during index construction.

	On low-dimensional datasets, DSG and WinFilter exhibit higher peak memory usage compared to other algorithms. This suggests that their index structures suffer from severe structural redundancy and are difficult to compress in low-dimensional scenarios.
	%因为低维空间中数据点之间的欧氏距离普遍较小，点与点之间更易互为近邻，从而导致图索引中产生大量冗余的边连接。

	%As Figure~\ref{fig:rangeFilter_memory_mb} shows, Faiss uses the least memory on low-dimensional datasets (Deep, YT-Audio). It stores minimal temporary data (e.g., vectors, centroids, distances) during index construction.

	For high-dimensional dataset WIT, iRange achieves the lowest peak memory usage. It pre-builds graphs only for a limited number of intervals, avoiding explicit index construction for all possible query ranges. Its index structure remains stable in high dimensions without significant graph growth. This effectively controls index size, making iRange more advantageous for high-dimensional datasets.
	In contrast, \textcolor{orange}{Milvus exhibits the highest memory consumption,}
	%Unify exhibits the highest memory consumption, indicating that its index construction process is highly sensitive to the dimensionality of the data.

		\textit{\textbf{\textcolor{orange}{Search peak memory.}}}

	%	\textcolor{orange}{As shown in Figure~\ref{fig:rangeFilter_search_memory_mb}, Faiss and Milvus have the lowest peak memory usage during search. Similar to attribute filtering, Faiss also has a smaller index size and no other additional memory usage, so the peak memory usage in range queries is slightly higher than the index size. Milvus reason:. WinFilter achieves the highest peak memory usage on almost all datasets. The reason is:. For the high-dimensional dataset WIT, UNIFY occupies the highest peak memory. The reason is:.}

		\textcolor{orange}{As for the search peak memory usage, we can find from the Figure ~\ref{fig:rangeFilter_search_memory_mb} that it is basically consistent with the build peak memory usage trend. The algorithm needs to load the complete index into memory during search. Therefore, algorithms with high building overhead, such as WinFilter, also have high search overhead; algorithms with low building overhead, such as Faiss, also have low search overhead. For Milvus, search does not require full index loading and there is additional metadata overhead when building, so the performance is different from the trend.}

@@

			\textbf{Large-Scale Dataset} & H & H & L & L &\textbf{H } &H  &M  &H  &M  &H  &\textbf{H}  &M/L  &L/M  &L/L  &M/M  &-  \\
			\textbf{Rare Attribute \& Low Range} &M  &M  &H  &H &\textbf{H}  &H  &H  &H  &L  &M  &\textbf{H} &M/M  &M/M  &L/L  &L/L  &L/L  \\
			\textbf{Robust to MultiModal Dataset} &H  &H  &L  &H  &H  &L  &M  &H  &M  &H  &H  &H/H  &M/M  &L/L  &L/L  &-/-  \\
			\textbf{Robust to Attribute Distribution} &L  &L  &H  &H  &H  &H  &-  &-  &-  &-  &-  &H/-  &L/-  &H/-  &L/-  &L/-  \\
			\textbf{High LID Dataset} &  &  &  &  &  &  &  &  &  &  &  &  &M  &L  &L/  &L/  \\
			\textbf{General-Purpose} &\faCheck  &\faCheck  &\faTimes  &\faTimes  &\faCheck  &\faTimes  &\faTimes  &\faCheck &\faTimes  & \faTimes &\faCheck&\faTimes  &\faCheck  &\faCheck  &\faTimes &\faTimes  \\
			\hline
		\end{tabular}
	}
	\centering
\end{table*}

>>>>

	%Similar to the evaluation of \textit{Attribute Filtering} algorithms, 
	Similarly, we also evaluate \textcolor{orange}{4} key metrics mentioned above in \textit{Range Filtering} algorithms. 
%	\textcolor{orange}{For the YT-Audio dataset with synthetic attributes and the YT-Audio dataset with real attributes, we can found that the data for index construction is basically the same, indicating that the authenticity of the attribute has almost no effect on index construction.However, the WIT dataset with synthetic labels is 2048-dimensional, while the WIT-Real dataset with real labels is 512-dimensional, so the dimension will cause a big difference.}

@@

	\textcolor{orange}{Note: For large-scale datasets, all algorithms are built using 64 threads for efficiency, except DSG, which only supports single-threaded construction. As a result, DSG takes significantly longer to build on Deep100M.
	}

	\textit{\textbf{Index size.}} 
	As shown in Figure~\ref{fig:rangeFilter_index_size_mb}, Among all algorithms, Faiss has the smallest index size. Because IVF only adds centroid data and partition information to the vector dataset, the IVF index is only slightly larger than the vector dataset.
	%On low-dimensional datasets (Deep, YT-Audio), 
	PASE requires the most storage. Its implementation likely wastes space due to alignment padding and inefficient encoding. %(e.g., Base64).
	\textcolor{orange}{DSG is an improvement over SeRF. Compared to SeRF, DSG stores additional information for each edge in its index structure, resulting in a significant increase in index size on large-scale datasets.
	}

	%SeRF maintains a compact index for low-dimensional data. However, for high-dimensional dataset (WIT), SeRF stores more edges and dynamic range data, significantly increasing its size.

	\textit{\textbf{Build peak memory.}}
%	As Figure~\ref{fig:rangeFilter_build_memory_mb} shows, Faiss uses the least memory on low-dimensional datasets. It stores minimal temporary data during index construction.
%	
%	On low-dimensional datasets, DSG and WinFilter exhibit higher peak memory usage compared to other algorithms. This suggests that their index structures suffer from severe structural redundancy and are difficult to compress in low-dimensional scenarios.
%	
%	For high-dimensional dataset WIT, iRange achieves the lowest peak memory usage. It pre-builds graphs only for a limited number of intervals, avoiding explicit index construction for all possible query ranges. Its index structure remains stable in high dimensions without significant graph growth. This effectively controls index size, making iRange more advantageous for high-dimensional datasets.
%	In contrast, \textcolor{orange}{Milvus exhibits the highest memory consumption,}

	As Figure~\ref{fig:rangeFilter_build_memory_mb} shows,

		\textit{\textbf{\textcolor{orange}{Search peak memory.}}}
		\textcolor{orange}{As for the search peak memory usage, we can find from the Figure ~\ref{fig:rangeFilter_search_memory_mb} that it is basically consistent with the build peak memory usage trend. As we analyzed in the section on attribute filtering.
		}

@@

			\textbf{Large-Scale Dataset} & H & H & L & L &\textbf{H } &H  &M  &H  &M  &H  &\textbf{H}  &M/L  &L/M  &L/L  &M/M  &-  \\
			\textbf{Rare Attribute \& Low Range} &M  &M  &H  &H &\textbf{H}  &H  &H  &H  &L  &M  &\textbf{H} &M/M  &M/M  &L/L  &L/L  &L/L  \\
			\textbf{Robust to MultiModal Dataset} &H  &H  &L  &H  &H  &L  &M  &H  &M  &H  &H  &H/H  &M/M  &L/L  &L/L  &-/-  \\
			\textbf{Robust to Attribute Distribution} &L  &L  &H  &H  &H  &H  &-  &-  &-  &-  &-  &H/-  &L/-  &H/-  &L/-  &L/-  \\
			\textbf{High LID Dataset} &H  &M  &H  &M  &H  &M  &M  &H  &M  &H  &H  &L/L  &M/H  &L/L  &L/M  &L/L  \\
			\textbf{General-Purpose} &\faCheck  &\faCheck  &\faTimes  &\faTimes  &\faCheck  &\faTimes  &\faTimes  &\faCheck &\faTimes  & \faTimes &\faCheck&\faTimes  &\faCheck  &\faCheck  &\faTimes &\faTimes  \\
			\hline
		\end{tabular}
	}
	\centering
\end{table*}
